---
title: "corrections"
author: "Tony Ni"
date: "6/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in Data

```{r}
setwd("~/harvard-summer-biostats")
df <- read_csv("data/wide_illinois.csv") #read in data
df1 <- na.omit(df) #get rid of na's
```

## Correcting upgradient values

Another thing we could think about is once we've classified wells into contaminated/not contaminated, we can use empirical distribution of the chemicals in the non-contaminated upgradient wells to try to figure out what a reasonable/standard concentration for each chemical to correct the contaminated wells.

Making same histograms (but leaving our contaminated wells) to see what the natural wells look like so that we can start to correct the contaminated values.

bootstrapping (take draw from noncontamianted wells and plug that in as if it were the value in the contaminated well, do that a bunch of times to get the, plug it into contamianted well value, take the difference btwn corrected upgradient and downgradient well values to use that to create a point estimate and a CI for the difference in the upgradient and downgradient values that should be corrected)



