---
title: "Exploratory Analysis"
author: "Tony Ni, Antonella Basso, Jose Lopez"
date: "6/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r, include = FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
```

Goals:

comparing how upgradients look vs downgradients through a graph (make sure the average thing/wrangling to fix it)

pick a chemical of interest

histograms of concentrations (one for upgradient vs downgradient for a single site)

compare the compare the 2 powerplants overall (upgradient at one vs the other)

regression

## Reading in Data

```{r}
setwd("~/harvard-summer-biostats")
illinois <- read_csv("data/illinois.csv") #read in data
```

## Checking unique/distinct levels

```{r}
#check how many observations there
nrow(illinois)

#checking distincts sites
unique(illinois$site)

#checking distinct gradient
unique(illinois$gradient)

#checking distinct gradient
unique(illinois$measurement.unit)

#checking disposal area
unique(illinois$disposal.area)

#checking type (L = land fuel, SI = surface impoundement, M = mixed/multiunit)
#there will be multiple different dumpsites at a powerplant site. some sites are wet (ponds of water where they throw ash of water), some sits are dry landfill (heaps of coal ash), the type is just a 
unique(illinois$type)

illinois %>%
  group_by(type) %>%
  summarize(n())

illinois %>%
  group_by(measurement.unit) %>%
  summarize(n())

#checking wells
illinois %>%
  group_by(well.id) %>%
  summarize(n())
```

Tony: Molybdenum to Thallium

use illinois and newyork

lower detection limit (if a chemical is really low, it can't differentiate beyond a certain low level, it might be the values with all teh same values)

all of the contaminants are harmful, some chemicals are more useful for detecting coal contamination (most important/indicative of coal contamination is boron (occurs at really high levels in coal)),

looking at the report given to us might be good about the different chemicals in coal (if we are curious)

total dissolved solids are probably not important, imprevise measurement of how much "crap is in the water" - doesnt represent any toxic chemicals

## Wrangling

```{r, warning = FALSE}
illinois1 <- illinois %>%
  select(site, disposal.area, type, well.id, gradient, contaminant, 
         measurement.unit, concentration) %>%
  mutate(well.id_contaminant = paste0(well.id, "_", contaminant)) %>% #for future use
  rename(c("disposal_area" = "disposal.area", "well_id" = "well.id",
           "unit" = "measurement.unit"))

#fixing 'contaminant' string by removing everything after the comma
illinois1$contaminant=gsub(", total", "", illinois1$contaminant)

#testing
avg_contaminant <- illinois1 %>%
  group_by(well_id, contaminant) %>% 
  summarise_each(funs(mean)) %>%
  select(1,2,8) #selecting only numeric columns

#temporarily uniting columns for joining in next step
temp <- avg_contaminant %>%
  unite("well.id_contaminant", well_id, contaminant)

#joining orig dataframe and avg_contaminant dataframe
combined <- left_join(temp, illinois1, by = "well.id_contaminant") %>%
  distinct(well.id_contaminant, .keep_all = TRUE) %>%
  separate(well.id_contaminant, c('well_id', 'contaminant'), sep="_") %>%
  select(1:8)

#spreading to wide data frame format to add missing info
combined2 <- combined %>% #collapse empty rows
  spread(contaminant, concentration.x) %>%
  group_by(well_id) %>%
  summarise_each(funs(first(.[!is.na(.)]))) %>%
  select(-c(unit))

#gathering back to long data frame format
combined3 <- combined2 %>%
  gather("contaminant", "concentration", 6:26)
```

## Summary Statistics

```{r}
illinois %>%
  group_by(contaminant) %>%
  summarize(mean_concentration = mean(concentration))
```

## Visualizations

```{r}
#looking at one specific well and the contaminant concentrations within it

ggplot(data = combined3 %>%
         filter(well_id == "03R"), aes(x = contaminant, y = concentration,
                                       fill = contaminant)) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  xlab("Contaminant") +
  ylab("Concentration") +
  ggtitle("Concentration of Contaminants in Well 03R") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#looking at one contaminant for all wells (counts)

ggplot(data = combined3 %>%
         filter(contaminant == "Antimony"), 
       aes(x = well_id, y = concentration, fill = well_id)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  xlab("Well ID") +
  ylab("Concentration") +
  ggtitle("Concentration of Antimony across Wells") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

The histogram is incompehensible, there are way too many wells in Illinois. 
Unsure, but having to make all these plots constantly seems unreasonable for all of these potential wells and contaminants to look through, maybe choose only those wells with concentration values greater than some certain value? (0.001 maybe?)

```{r}
ggplot(data = combined3 %>%
         filter(contaminant == "Antimony"),
       aes(x = well_id, y = concentration, fill = well_id)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  xlab("Well ID") +
  ylab("Concentration") +
  ggtitle("Concentration of Antimony across Wells") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

levels(as.factor(combined3$well_id))
```

#Comparing Upgradient Wells

```{r}
#get the first 10 highest avg mg/l (not ph or total sulfate)
#not pH, total dissolve solids, or radium b/c different units

dont_use <- c("pH", "Radium 226+228", "Total Dissolved Solids")

#pull out the names (from top 10 high to low avg concentrations) of contaminants
chemicals <- combined3 %>%
	filter(contaminant != "pH") %>% #take our nonuseful contaminants
  filter(contaminant != "Radium 226+228") %>%
  filter(contaminant != "Total Dissolved Solids") %>%
	group_by(contaminant) %>% 
	summarise(avg_conc = mean(concentration, na.rm = TRUE)) %>%
	arrange(desc(avg_conc)) %>%
  slice(1:10) %>% #there only seem to be 5 chemicals of nonzero(basically) values
	pull(contaminant)

chemicals

combined3 %>%
	group_by(contaminant) %>% 
	summarise(avg_conc = mean(concentration))

```

Is doing the mean of concentrations amongst all wells OK to do? Some wells like arsenic have wells with 0.001 (sign that measurement was insignificant/below threshold), and taking the avg sometimes makes it return NA as the mean (I THINK??? what does NA mean...?) ASK LULI (i think some of the chemicals in the paper had thresholds of DANGER that were in mg and others in microg, how would we account for this...?)

>

https://postimg.cc/ZCc072JM

```{r}
#plot in a faceted grid bar plots of concentration vs. contaminant for all wells
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant %in% chemicals), 
      aes(x = contaminant, y = concentration)) + 
      geom_bar(stat = "Identity") +
      facet_wrap(well_id ~ .) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

The concentration of calcium is so high in some of the wells, it makes it difficult to look at the other contaminants' concentration across all wells, perhaps we should make different facetted plots for all contaminants.

```{r}
#Boron
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant %in% chemicals) %>%
        filter(contaminant == "Boron"), 
      aes(x = contaminant, y = concentration)) + 
      geom_bar(stat = "Identity") +
      facet_wrap(well_id ~ .) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
#Calcium
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant %in% chemicals) %>%
        filter(contaminant == "Calcium"), 
      aes(x = contaminant, y = concentration)) + 
      geom_bar(stat = "Identity") +
      facet_wrap(well_id ~ .) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
#Chloride
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant %in% chemicals) %>%
        filter(contaminant == "Chloride"), 
      aes(x = contaminant, y = concentration)) + 
      geom_bar(stat = "Identity") +
      facet_wrap(well_id ~ .) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
#Fluoride
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant %in% chemicals) %>%
        filter(contaminant == "Fluoride"), 
      aes(x = contaminant, y = concentration)) + 
      geom_bar(stat = "Identity") +
      facet_wrap(well_id ~ .) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Let's make a histogram to show distribution of certain contaminants:

```{r}
#Molybdenum
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "Molybdenum"), 
      aes(x = concentration)) + 
      geom_histogram() +
      geom_vline(xintercept = 40/1000) + #threshold value
  xlab("Concentration (mg/L)") +
  ylab("Count") +
  ggtitle("Distribution of Molybdenum amongst all Upgradient Wells") +
  theme_bw()

#pH
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "pH"), 
      aes(x = concentration)) + 
      geom_histogram() +
  xlab("units") +
  ylab("Count") +
  ggtitle("Distribution of pH amongst all Upgradient Wells") +
  theme_bw()

#Radium 226+228
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "Radium 226+228"), 
      aes(x = concentration)) + 
      geom_histogram() +
      geom_vline(xintercept = 5) + #threshold value
  xlab("Concentration (pCi/L)") +
  ylab("Count") +
  ggtitle("Distribution of Radium 226+228 amongst all Upgradient Wells") +
  theme_bw()

#Selenium
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "Selenium"), 
      aes(x = concentration)) + 
      geom_histogram() +
      geom_vline(xintercept = 50/1000) + #threshold value
  xlab("Concentration (mg/L)") +
  ylab("Count") +
  ggtitle("Distribution of Selenium amongst all Upgradient Wells") +
  theme_bw()

#Sulfate
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "Sulfate"), 
      aes(x = concentration)) + 
      geom_histogram() +
      geom_vline(xintercept = 500) + #threshold value
  xlab("Concentration (mg/L)") +
  ylab("Count") +
  ggtitle("Distribution of Sulfate amongst all Upgradient Wells") +
  theme_bw()

#Thallium
ggplot(data = combined3 %>%
        filter(gradient == "Upgradient") %>%
        filter(contaminant == "Thallium"), 
      aes(x = concentration)) + 
      geom_histogram() +
      geom_vline(xintercept = 2/1000) + #threshold value
  xlab("Concentration (mg/L)") +
  ylab("Count") +
  ggtitle("Distribution of Thallium amongst all Upgradient Wells") +
  theme_bw()
```

We can see that there are some wells with concentrations that lie far beyond the threshold value (obtained from the PDF), and it would be nice to identify them. We will pull out the observatiosn with concentrations past the threshold and place them within a new dataset.

```{r}
#creating vector of contaminants
contaminant <- unique(combined3$contaminant)

#creating vector of threshold values for each contaminant
threshold <- c(6/1000, 10/1000, 2, 4/1000, 3, 5/1000, NA, NA, 100/1000, 
               6/1000, NA, 15/1000, 40/1000, 2/1000, 40/1000, NA, 5, 
               50/1000, 500, 2/1000, NA)

contam_t <- cbind(contaminants, threshold) %>%
  na.omit()

#creating function to obtain all observations with values above threshold
getOverThreshold <- function(df){
  datalist = list()
  for(i in 1:nrow(contam_t)){ #for each contaminant i
    df1 <- filter(df, gradient == "Upgradient")
    df2 <- filter(df1, contaminant == contam_t[i])
    data <- filter(df2, concentration > contam_t[nrow(contam_t) + i])
    datalist[[i]] <- data
  }
  toReturn <- do.call(rbind, datalist)
  
  return(toReturn)
}

overthreshold_df <- getOverThreshold(combined3) #df with all observations over threshold value
```


How do we determine what is the NORM? What would the usual distribution of sulfate be if it were not contaminated?

> Less contaminated upgradient wells probably have less amts of chemicals to some degree -- some wells are probably easier to notice that are less contaminated than others

------------------------------------------

Comments/Notes Below:

we could see that in an exploratory analysis, look at histogram of different chemcials at each well -- there should be variation btwn upgradients at the same sites and general amt of different between upgradient at one site vs another site... one source of variation is the site and wells will also have some variation in their measurement too... make histograms for all chemcials and visualize them side by side and see if theres any differnces between upgradients and downgradients, and check if any differences between upgradients.. statistically there is a way... group them by upgradient vs downgradient (groupby chemicals and plot x variable concentration, and facet by the gradient (either up or down) and get histograms for each)

In the powerpoint, it talks about comparing upgradient wells within 100 mi radius of each other, how should we go about getting this distance data since it isn't provided? Should we try to use outside data/geospatial data to do this?

> For now, looking at data w/o distance probably is fine -- we might incorporate distance if we get further/have time.

Are all the contaminants which are being measured of danger? Or are there only certain ones that should be looked at which are dangerous?

>

I realized a lot of the values are inconsistent in their units, some are ug/L others of mg/L, etc. It would be nice to standardize these across the board in the pre-wrangling stage...

> http://www.eai-labs.com/assets/docs/radioactive_in_water.pdf says "micrograms per liter (ug/L) can be converted to pCi/L by multiplying the U (mg/L) by 0.67"

Plot 1: Compare concentrations at a small number of upgradient wells with other upgradient wells at the same site.

Looking at https://www.uky.edu/KGS/coal/coal-major-minor-trace-elements.php , we find that the major elements in coal consist of: Hydrogen, Carbon, Nitrogen, Oxygen, and Sulfur with the minor elements in coal being: Sodium, Aluminum, Silicon, Phosphorous, Potassium, Calcium, Titanium, Manganese, and Iron

The contaminants being measured in our case are Sulfate (a compound containing sulfur), Calcium, 




