---
title: "Exploratory Analysis"
author: "Tony Ni, Antonella Basso, Jose Lopez"
date: "6/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
library(mdsr)
library(tidyverse)
library(ggplot2)
```

This code chunk below does things!!!

Tony typed this here.

Nelly is a baddie.

Hello <3


comparing how upgradients look vs downgradients through a graph (make sure the average thing/wrangling to fix it)

pick a chemical of interest

histograms of concentrations (one for upgradient vs downgradient for a single site)

compare the compare the 2 powerplants overall (upgradient at one vs the other)

regression

## Reading in Files

First, we need to read in our data.

```{r}
new_york <- read_csv("data/new_york.csv") #read in data

#check how many observations there
nrow(new_york)

#checking distincts sites
unique(new_york$site)

#checking distinct gradient
unique(new_york$gradient)

#checking distinct gradient
unique(new_york$measurement.unit)

#checking disposal area
unique(new_york$disposal.area)

#checking type (L = land fuel, SI = surface impoundement, M = mixed/multiunit)
#nelly asks what does type mean???
unique(new_york$type)

new_york %>%
  group_by(type) %>%
  summarize(n())

#ask if we might want to remove the other measurements for consistency?????
#nelly asks can we find a way to convert these easily!

new_york %>%
  group_by(measurement.unit) %>%
  summarize(n())

#checking wells
new_york %>%
  group_by(well.id) %>%
  summarize(n())
```

## Wrangling

```{r}
new_york1 <- new_york %>%
  select(site, disposal.area, type, well.id, gradient, contaminant, 
         measurement.unit, concentration) %>%
  filter(measurement.unit == "mg/l")  %>% #for now, filtering out non mg/l observations
  mutate(well.id_contaminant = paste0(well.id, "_", contaminant)) #for future use

#testing
avg_contaminant <- new_york1 %>%
  group_by(well.id, contaminant) %>% 
  summarise_each(funs(mean)) %>%
  select(1,2,8) #selecting only numeric columns

#temporarily uniting columns for joining in next step
temp <- avg_contaminant %>%
  unite("well.id_contaminant", well.id, contaminant)

#joining orig dataframe and avg_contaminant dataframe
combined <- left_join(temp, new_york1, by = "well.id_contaminant") %>%
  distinct(well.id_contaminant, .keep_all = TRUE) %>%
  separate(well.id_contaminant, c('well.id', 'contaminant'), sep="_") %>%
  select(1:8)
```

## Summary Statistics

```{r}
new_york %>%
  group_by(contaminant) %>%
  summarize()
```


```{r}
object <- new_york %>%
  group_by(contaminant) %>%
  summarize(mean_concentration = mean(concentration))

object
```

## Visualizations

```{r}
#ggplot(name of dataset, aes(x = variable)) + geom_whatever

#first histogram (really skewed)
ggplot(new_york, aes(x = concentration)) + geom_histogram()
```

## Wrangling



```{r}
antimony <- 
```

